% FREUID Challenge 2026 — Technical Report
% Compile: pdflatex freuid_technical_report.tex && bibtex freuid_technical_report && pdflatex x2
% License: CC0 template (organizer-provided) — content below is our own.

\documentclass[11pt,a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}

\geometry{margin=2.5cm}
\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  citecolor=blue!60!black,
  urlcolor=blue!60!black,
}

\newcommand{\teamname}{Nadhir Hasan}
\newcommand{\kaggleteam}{Nadhir Hasan}
\newcommand{\contactemail}{hasannadhir2003@gmail.com}
\newcommand{\reportdate}{\today}
\newcommand{\commitsha}{0c034d5394794fb0e6b4c237e5c8fd536013b90c}

\title{%
  \textbf{The FREUID Challenge 2026}\\[0.4em]
  \large Technical Report --- \teamname
}
\author{%
  Nadhir Hasan\\
  \small
  \texttt{\contactemail}
}
\date{Report date: \reportdate}

\begin{document}
\maketitle

\begin{abstract}
We present our solution to the FREUID Challenge 2026 identity-document fraud
detection task. Both of our selected final submissions use the same architecture
--- a DINOv2-L Vision Transformer with low-rank adapters (LoRA, rank 16) and a
per-patch multiple-instance-learning (MIL) classification head at $448\times728$
resolution --- trained with an annotation-driven synthetic fraud generator
(portrait swap, cross-card region swap, text-field editing, erase-and-retype)
built on manually labeled per-type face and text-field bounding boxes. The two
picks use \emph{identical} inference-time options and differ only in training
data: \textbf{Pick 1} trains on FREUID only and maximizes in-domain performance
(public leaderboard 0.00060); \textbf{Pick 2} trains on 100\% of FREUID plus
80{,}000 images from the externally licensed IDNet dataset, trading public score
(0.00191) for a measured order-of-magnitude improvement in cross-document-type
robustness (FREUID Score 0.0116 vs.\ chance-level for FREUID-only models on
35{,}874 held-out real-world documents). Both picks share two inference-time
enhancements selected on external validation data and confirmed by isolated
public-leaderboard probes: a re-aggregation of the per-patch MIL head (mean of
top-5\% patch logits, weight 0.25, blended with the attention branch) and
4-scale logit-averaged test-time augmentation --- each improved every model it
was applied to, and they compounded. We document why
score calibration is provably a no-op for the FREUID metric, and why we selected
checkpoints on held-out external data rather than in-domain validation. Code,
weights, and a network-isolated Docker container reproducing both picks from one
image are provided in the linked repository.
\end{abstract}

\noindent\textbf{Competition:} The FREUID Challenge 2026 (IJCAI-ECAI 2026)\\
\textbf{Kaggle team:} \kaggleteam\\
\textbf{Code:} \url{https://github.com/nadhirhasan/FREUID-Challenge-2026} (commit \texttt{\commitsha})\\

\section{Introduction}
\label{sec:intro}
Identity-document fraud detection must generalize across a threat model that spans
physical tampering, GenAI-driven digital manipulation, and print-and-capture
(``analog hole'') attacks, and across document types and issuing countries that are
not exhaustively represented in any single training set. The FREUID Challenge
formalizes this with a training set covering 5 document types and a scoring metric
(FREUID Score, Section~\ref{sec:results}) that combines a global ranking measure
(AuDET) with a strict low-false-alarm operating point (APCER@1\%BPCER), penalizing
solutions that overfit to easy global separability while failing at the
operating point that matters for production KYC systems~\cite{freuid2026}.

Our solution strategy is built around a specific empirical finding from our own
development process: a classifier trained \emph{only} on the FREUID distribution
can reach extremely low FREUID Score on the public leaderboard while being
statistically indistinguishable from random (AuDET $\approx 0.47$--$0.50$) on
real-world identity documents it has never seen
(Section~\ref{sec:validation}). We further measured, by nearest-template
classification of the public test images, that the test set is not purely
FREUID-like: it contains at least one document type absent from training (a
Liberia voter card) and a tail of physically captured documents (cards
photographed on desks, including a physical photo-substitution attack), whereas
the training set is 99.97\% digital renders. Because the private leaderboard is
$\approx$95\% of the test data, and because a $\sim$1\% subpopulation of
domain-shifted attacks can dominate the missed-attack count at sub-1\% APCER
levels, we selected \emph{two} complementary final submissions: the strongest
in-domain model, and a robustness hedge trained additionally on external
real-document data.

\section{Method}
\label{sec:method}

\subsection{Overview}
Each input image is loaded at full resolution, converted to RGB, letterboxed
(aspect-ratio-preserving resize + zero-padding) to a fixed $448\times728$ canvas,
and normalized with standard ImageNet statistics. The backbone produces a token
sequence; a per-patch head scores each spatial token independently and aggregates
into a single image-level fraud logit (aggregation details in
Section~\ref{sec:postproc}). A sigmoid converts this logit to the submitted score
in $[0,1]$; no calibration is applied (Section~\ref{sec:calibration}).

\subsection{Model architecture}
Backbone: \texttt{vit\_large\_patch14\_reg4\_dinov2} (DINOv2-L~\cite{dinov2} with
register tokens), loaded via \texttt{timm}, Apache-2.0 licensed pre-trained weights, 24
transformer blocks, embedding dimension 1024. We attach a from-scratch, minimal
LoRA implementation (rank 16, $\alpha=32$, dropout 0.05; no third-party
\texttt{peft} dependency) to the \texttt{qkv}, \texttt{proj}, \texttt{fc1}, and
\texttt{fc2} linear layers of every transformer block; the backbone itself is
frozen. On top of the token sequence we place a lightweight per-patch
multiple-instance-learning (MIL) head: a 2-layer MLP (hidden size 512, GELU,
dropout 0.3) scores every non-prefix token; during training the image-level
logit is the equal-weight mean of (i) the top-10\% most suspicious patch logits
and (ii) an attention-weighted sum over all patch logits. This design was chosen
to push the model toward \emph{local}, document-type-agnostic forgery cues (blend
seams, splice boundaries, compression discontinuities) rather than global,
country-specific layout statistics. Trainable parameters: LoRA adapters + head
only ($\approx$7M of $\approx$304M total).

\subsection{Training objective and procedure}
Loss: Focal BCE ($\alpha=0.25$, $\gamma=2.0$) on the whole-image logit.
Optimizer: AdamW, two parameter groups (head lr $1\times10^{-3}$, LoRA lr
$2\times10^{-4}$, weight decay 0.05/0 respectively), cosine learning-rate
schedule with 5\% linear warmup, mixed precision (fp16). Both picks were trained
on a single NVIDIA RTX A4500 (20GB).

\textbf{Pick 1 (\texttt{cv3}).} FREUID only: fold 0 of a stratified 5-fold split
($\approx$55{,}500 training images/epoch), 3 scheduled epochs; the submitted
checkpoint is epoch 1, which had the best public-leaderboard probe.

\textbf{Pick 2 (\texttt{cv5}, epoch 2).} 100\% of the FREUID training set
(69{,}352 images) plus 80{,}000 balanced IDNet images spanning all 10 available
countries ($\approx$149{,}000 images/epoch), 5 scheduled epochs with per-epoch
validation on 4{,}000 IDNet images sampled exclusively from pool rows \emph{not}
selected into the 80k training cap (same countries, zero row-level overlap).
Epoch 2 was selected: its successor (epoch 3) improved the IDNet validation score
(0.0105 vs.\ 0.0166) but scored \emph{worse} on a public-leaderboard probe (0.007
vs.\ 0.0035), a reversal consistent with our finding that no single local metric
reliably ranks checkpoints (Section~\ref{sec:validation}); training was stopped
at that point.

\textbf{Synthetic fraud generation.} 25\% of genuine training images (from both
FREUID and IDNet sources) are converted to synthetic fraud examples each epoch,
using one of four attack types, dispatched via manually annotated per-document-type
face and text-field bounding boxes (Section~\ref{sec:external-data}):
cross-card region/portrait swap (donor of the same document type), text-field
edit (same-line content shifted and re-blended), erase-and-retype (ink removed via
inpainting, new text rendered with a system font), and a generic self-blend
fallback. Blending uses a mix of feathered alpha compositing and seamless
(Poisson) cloning to avoid a single, learnable seam signature. This attack suite,
compared against a simpler untargeted self-blend-only baseline under an otherwise
identical recipe, improved the public leaderboard score from 0.00169 to 0.00096
in isolated ablation.

\subsection{Inference-time enhancements (post-processing)}
\label{sec:postproc}
The FREUID metric depends only on the score \emph{ranking} (AuDET is rank-based;
the APCER threshold is a quantile of the model's own genuine scores), so the only
admissible post-processing levers are rank-changing ones. We evaluated two,
selecting configurations on held-out external data before confirming with single
public-leaderboard probes:

\begin{itemize}[noitemsep]
  \item \textbf{Patch re-aggregation (both picks).} At inference we recompute the
    image logit from the per-patch logits as
    $0.25\cdot\mathrm{mean}(\text{top-}5\%\ \text{patch logits}) +
    0.75\cdot(\text{attention branch})$, replacing the training-time top-10\%/0.5
    blend. A grid over top-$k$ fraction $\times$ branch weight on 12{,}000
    held-out IDNet images selected this configuration; it subsequently improved
    the public score of \emph{every} model it was applied to (Pick 1 base:
    0.00096 $\rightarrow$ 0.00086; Pick 2 base: 0.0035 $\rightarrow$ 0.00263). We
    attribute the gain to softened single-patch false-positive spikes on genuine
    documents, which lower the 1\%-BPCER threshold --- precisely the metric's
    sensitive component at our operating level.
  \item \textbf{Multi-scale TTA (both picks).} Logit-average over input scales
    $\{0.85, 0.9, 1.0, 1.1\}$ (each rounded down to the ViT patch grid). Selected
    from all 31 scale combinations on the 12k external subsample and confirmed on
    the full 35{,}874-image pool (0.0154 $\rightarrow$ 0.0116, $-$25\%). Our prior
    --- informed by the image-forensics literature's warning that resampling
    destroys low-level artifact signals~\cite{deepid2025}, and by TTA alone
    slightly worsening Pick 2's public score (0.0035 $\rightarrow$ 0.00369) ---
    was to withhold TTA from the FREUID-only Pick 1; a direct leaderboard probe
    refuted that prior: combined with patch re-aggregation, TTA improved Pick 1
    from 0.00086 to 0.00060 and Pick 2 from 0.00263 to 0.00191. The two levers
    compounded on both models, so both picks ship with both levers.
\end{itemize}

\subsection{Score calibration}
\label{sec:calibration}
No score calibration or thresholding is applied: the submitted label is
$\sigma(\text{logit})\in(0,1)$ directly. The FREUID Score is invariant to any
strictly monotone transform of the scores (both AuDET and the
quantile-thresholded APCER depend only on ranking) --- we verified this
numerically against our exact metric implementation --- so temperature scaling,
normalization, or clipping are provably no-ops (clipping is actively harmful, as
score ties at the clip boundary distort the genuine-score quantile).

\section{Data}
\label{sec:data}

\subsection{FREUID training set}
\label{sec:external-data}
We use the official \texttt{train/} images and \texttt{train\_labels.csv}
metadata. Pick 1 trains on fold 0 of a stratified 5-fold split by
(\texttt{type}$\times$\texttt{label}); Pick 2 trains on 100\% of the training
set. All 5 FREUID document types are present in training in both cases. We
hand-annotated portrait and text-field bounding boxes for every document type
(5 FREUID + 10 IDNet countries) to drive the synthetic attack generator
precisely rather than at random image locations; these annotations are included
in the repository.

\subsection{External data and pre-trained models}
\begin{itemize}[noitemsep]
  \item \textbf{DINOv2}~\cite{dinov2} pre-trained ViT-L/14 weights, via
    \texttt{timm} (Apache-2.0 license). Used by both picks.
  \item \textbf{IDNet}~\cite{idnet2024} (synthetically generated but
    realistically captured identity documents across 10 European countries),
    CC BY 4.0 license, publicly available. Used by Pick 2 only: 80{,}000
    balanced images across all 10 countries mixed into training; 4{,}000
    disjoint images (same countries, zero row overlap) used for per-epoch
    validation. For the \emph{external generalization benchmark} reported
    throughout (Section~\ref{sec:validation}) we evaluate on the full
    35{,}874-image pool of two countries (Estonia, Slovakia); for Pick 1 and
    other FREUID-only models this is a fully blind test (those models never saw
    IDNet in any form), while for Pick 2 it is partially optimistic (disjoint
    images, but the countries appear in training) --- we state this caveat
    wherever the number is used.
\end{itemize}
Both artifacts are publicly available under licenses compatible with the
competition's external-data policy; no proprietary or non-freely-accessible data
was used. We deliberately avoided any model, checkpoint, or code released under
a non-OSI-approved or non-commercial license (e.g.\ TruFor, ConvNeXt V2, DINOv3)
anywhere in the training or inference pipeline.

\subsection{Validation strategy}
\label{sec:validation}
During development we established, through repeated controlled comparisons, that
in-domain FREUID validation (stratified disjoint folds, even with test-like
corruption applied) does not reliably predict public-leaderboard rank. We traced
part of this to visual asset/persona reuse within the FREUID training
distribution (near-duplicate faces and field values recur across nominally
disjoint splits). Our resulting protocol: every decision that could be validated
was scored with the exact competition metric on held-out IDNet data, and final
candidates were confirmed with single leaderboard probes.

The decisive experiment for the two-pick strategy evaluated all candidate
models, identically, on the full 35{,}874-image Estonia+Slovakia IDNet pool:
FREUID-only models are at chance there (Pick 1: FREUID Score 0.9727; a
full-data FREUID-only variant: 0.9424) despite public scores of 0.00096 and
0.00178 respectively, while Pick 2 reaches 0.0116 --- yet all these models
correlate at Pearson $\ge$0.994 on public-test predictions, i.e.\ they share
the same in-domain ranking and differ only in out-of-domain behavior. Pick 1's
in-domain generalization is separately supported by its genuinely held-out,
corruption-augmented 20\% FREUID fold (FREUID Score 0.00019). Private test
labels were never used, seen, or referenced during model selection.

\section{Inference}
\label{sec:inference}
The Docker entrypoint (\texttt{docker/prepare\_submission.py}) discovers every
image directly under \texttt{/data/}, decodes each image exactly once, produces
the configured scale tensors, and writes \texttt{/submissions/submission.csv}
with columns \texttt{id,label}. One image reproduces both picks via documented
environment variables (inference orchestration only; identical frozen weights):

\begin{itemize}[noitemsep]
  \item \textbf{Pick 1 (image defaults):} \texttt{FREUID\_MODEL\_PATH=/models/cv3\_fold0.pt},
    \texttt{FREUID\_PATCH\_AGG=0.05,0.25}, \texttt{FREUID\_TTA\_SCALES=0.85,0.9,1.0,1.1}.
  \item \textbf{Pick 2:} \texttt{-e FREUID\_MODEL\_PATH=/models/cv5\_full\_ep2.pt}
    (all other settings identical --- the model path is the only difference
    between the two picks).
\end{itemize}

The container requires no network access at runtime (\texttt{--network none});
all weights are baked in at build time. Measured cost of the shared 4-scale
configuration on an RTX A4500: $\approx$160\,ms/image (JPEG decoded once per
image, decode overlapped with GPU compute via threaded prefetch), i.e.\
$\approx$6.3\,h for the 142{,}818-image hidden set on our A4500 and
$\approx$2--3\,h on an A100 (2--3$\times$ the A4500's FP16 throughput and
memory bandwidth) --- within the 6-hour budget per pick, with batch size and
decode threads tunable via environment variables for further headroom.
Inference auto-detects CUDA and falls back to CPU if unavailable.

\section{Results}
\label{sec:results}

\begin{table}[h]
  \centering
  \caption{Main results (FREUID Score; lower is better). External = full
  35{,}874-image IDNet Estonia+Slovakia pool (blind for FREUID-only models;
  see Section~\ref{sec:external-data} for Pick 2's caveat).}
  \label{tab:results}
  \begin{tabular}{lccl}
    \toprule
    Model / configuration & Public LB $\downarrow$ & External $\downarrow$ & Notes \\
    \midrule
    Pick 1: \texttt{cv3} + re-agg + TTA4            & \textbf{0.00060} & ---    & Selected (primary) \\
    \quad \texttt{cv3} + re-agg only                & 0.00086 & ---    & \\
    \quad \texttt{cv3} raw                          & 0.00096 & 0.9727 & chance-level externally \\
    Pick 2: \texttt{cv5} ep2 + re-agg + TTA4        & 0.00191 & \textbf{0.0116}$^{*}$ & Selected (hedge) \\
    \quad \texttt{cv5} ep2 raw                      & 0.0035  & 0.0154 & \\
    \quad \texttt{cv5} ep2 + re-agg only            & 0.00263 & 0.0137$^{\dagger}$ & \\
    \quad \texttt{cv5} ep2 + TTA4 only              & 0.00369 & 0.0116 & \\
    FREUID-only, 100\% data (\texttt{cv6}, ep3)     & 0.00178 & 0.9424 & not selected \\
    FREUID+IDNet fold model (\texttt{cv4}, ep2)     & 0.00467 & 0.0933$^{\ddagger}$ & earlier iteration \\
    \bottomrule
  \end{tabular}

  \vspace{2pt}
  {\footnotesize $^{*}$full-pool confirmation of the externally selected config;
  $^{\dagger}$12k-subsample estimate; $^{\ddagger}$4{,}000-image sample of the
  same two countries, fully blind for this model (never trained on them).}
\end{table}

\noindent\textbf{Ablation --- attack realism.} Holding the architecture, data
split, and all other hyperparameters fixed, replacing an untargeted
self-blend-only synthetic-fraud generator with the annotation-driven attack
suite improved the public leaderboard score from 0.00169 to 0.00096.

\noindent\textbf{Ablation --- data mixing for generalization.} FREUID-only
models (fold-based or full-data) are at chance on external documents regardless
of their public score (Table~\ref{tab:results}); mixing IDNet into training is
the single lever that produced non-trivial external performance, at a public
cost that inference-time enhancements later recovered (0.0035 $\rightarrow$
0.00191).

\noindent\textbf{Ablation --- inference-time enhancements.} Patch
re-aggregation improved every model it was applied to (public: $-$10\% for
Pick 1 base, $-$25\% for Pick 2 base). Multi-scale TTA improved external
robustness by 25\% (full-pool confirmed) and, combined with re-aggregation,
compounded on the public leaderboard for both picks (Pick 1: 0.00086
$\rightarrow$ 0.00060; Pick 2: 0.00263 $\rightarrow$ 0.00191), despite TTA
alone slightly worsening Pick 2's public score --- each lever was probed in
isolation and in combination before selection.

\section{Reproducibility}
\label{sec:repro}
\begin{itemize}[noitemsep]
  \item \textbf{Repository:} \url{https://github.com/nadhirhasan/FREUID-Challenge-2026}, commit \texttt{\commitsha}.
  \item \textbf{One Docker image, both picks:} \texttt{docker build -t freuid-repro:local .}
    from the repository root; run commands for each pick (and the
    submission$\,\leftrightarrow\,$command$\,\leftrightarrow\,$checksum mapping)
    are documented in the repository \texttt{README.md}. Weights are baked in at
    \texttt{/models/cv3\_fold0.pt} and \texttt{/models/cv5\_full\_ep2.pt}.
  \item \textbf{Lean checkpoints:} the repository ships only the trained
    parameters (192 LoRA matrices + 8 head tensors, $\approx$27\,MB per pick);
    the frozen DINOv2-L backbone is restored from the official \texttt{timm}
    release at Docker \emph{build} time (runtime remains network-free). The
    conversion is lossless and audited by \texttt{src/make\_lean\_ckpt.py}: all
    343 backbone tensors of the full training checkpoints are bit-identical to
    the timm release, and lean-loaded models produce bit-identical fp32 logits
    to the full checkpoints.
  \item \textbf{Submitted CSVs (sha256):}
    \texttt{sub\_cv3\_pagg.csv} = \texttt{9f547e04\ldots{}aa376a6};
    \texttt{sub\_cv5\_full\_ep2\_pagg\_tta4.csv} = \texttt{91392d1c\ldots{}96ebd73}
    (full hashes in \texttt{README.md}). CSVs contain model predictions for the
    7{,}821 locally available public-test images and a 0.5 placeholder for
    hidden ids; bit-exact float reproduction across differing GPU hardware is
    not guaranteed, rank-identical scores are.
  \item \textbf{Dependencies:} Python 3.11, PyTorch 2.6.0 (+cu124), timm 1.0.27,
    opencv-python-headless 4.10.0.84 --- see \texttt{docker/requirements.txt}.
  \item \textbf{Randomness:} training uses unseeded per-worker RNGs for
    augmentation only; inference is deterministic given fixed weights.
  \item \textbf{Network:} inference requires \texttt{--network none}; no
    downloads occur at container runtime.
\end{itemize}

\section{Team and contributions}
Single-member team: Nadhir Hasan (all components --- data analysis, annotation,
training, validation protocol, inference, packaging).

\section*{Acknowledgments}
We thank the FREUID Challenge organizers for the FantasyID/FREUID dataset and
the IDNet authors for their permissively licensed dataset.

\bibliographystyle{plain}
\bibliography{references}

\appendix
\section{Hyperparameters}
\label{app:hparams}
\begin{table}[h]
  \centering
  \small
  \begin{tabular}{lll}
    \toprule
    Hyperparameter & Pick 1 (\texttt{cv3}) & Pick 2 (\texttt{cv5} ep2) \\
    \midrule
    Image size & \multicolumn{2}{c}{$448\times728$ (letterboxed)} \\
    Batch size (train) & \multicolumn{2}{c}{6 (accum 4, effective 24)} \\
    Learning rate (head / LoRA) & \multicolumn{2}{c}{$1\times10^{-3}$ / $2\times10^{-4}$} \\
    Weight decay (head / LoRA) & \multicolumn{2}{c}{0.05 / 0.0} \\
    LoRA rank / $\alpha$ / dropout & \multicolumn{2}{c}{16 / 32 / 0.05} \\
    Synthetic-attack rate & \multicolumn{2}{c}{0.25} \\
    FREUID training data & fold 0 ($\approx$80\%) & 100\% \\
    IDNet training data & none & 80{,}000 (10 countries, balanced) \\
    Scheduled epochs / submitted & 3 / epoch 1 & 5 (stopped after 3) / epoch 2 \\
    Patch re-agg (inference) & \multicolumn{2}{c}{top-5\% frac, branch weight 0.25} \\
    TTA scales (inference) & \multicolumn{2}{c}{0.85, 0.9, 1.0, 1.1 (logit-avg)} \\
    Seed (fold split) & \multicolumn{2}{c}{42} \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Additional ablations}
See Section~\ref{sec:results} for the controlled single-lever ablations (attack
realism; data mixing; inference-time enhancements) that produced the final
two-pick configuration.

\end{document}
